---
title: "Volleyball Percentiles"
---

We'll be calculating percentiles to see how good Maryland was in service aces in the 2024 season. We'll use `mutate()` and a new function `ntile()` to calculate percentiles. Get started by loading the tidyverse in a code block, then load the CSV file here: https://raw.githubusercontent.com/dwillis/NCAAWomensVolleyballData/refs/heads/main/data/ncaa_womens_volleyball_matchstats_2024.csv" and save it to a dataframe called `matches`.

### Task 1: Load the tidyverse
```{r}
library(tidyverse)
```

### Task 2: Load the data

```{r}
read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensVolleyballData/refs/heads/main/data/ncaa_womens_volleyball_matchstats_2024.csv")
```

### Task 3: Create totals of aces for each team using `group_by()` and `summarize()`

```{r}
vb_stats<-read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensVolleyballData/refs/heads/main/data/ncaa_womens_volleyball_matchstats_2024.csv")

team_aces <- vb_stats %>%
  group_by(team) %>% 
  summarize(total_aces = sum(aces, na.rm = TRUE))

```
Had to use chat Gpt for this one and what I've gathered in using it is that it created a chart using the vb_stats chart, and made a group with the teams and then summerized the total_aces by equalling total aces to the sum of aces for that team, which it knows to do bevause of the pipe operator  %>% .

### Task 4: Calculate percentiles using `mutate()` and `ntile()`

```{r}
  
team_aces <- team_aces |> mutate(ace_percentile = ntile(total_aces, 100))

```

Ok i used chat gpt again and the first recommendation was "mutate(ace_percentile = ntile(total_aces, 100))"


### Task 5: Describe the percentiles

Write a couple of sentences that puts Maryland into context with the other teams in the dataset. You can do some Internet research to help with this.

So as far as the volleyball team goes they rank in the 95th percentile. That means that they scored more aces than 95% of other teams, which is great. but unfortunately that doesnt transalte to over all wins, in which maryland falls to be ranked 152. 

When you are done, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.






