---
title: "ANOVA for Journalists: Analyzing College Admissions Test Scores"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Introduction

In this exercise, we'll explore a common journalism question: *Are there significant differences in standardized test scores across different types of colleges?*

We'll use ANOVA (Analysis of Variance) to determine if the observed differences in SAT scores across three college types (Public, Private, and Community College) represent real patterns or just random variation.

## Task 1: Load Libraries and Create Sample Data

```{r}
library(tidyverse)
```

```{r create-data}
# Set seed for reproducibility
set.seed(123)

# Create simulated SAT scores for admitted students across three college types
# Scale is 400-1600 for total SAT
public_college <- rnorm(100, mean = 1350, sd = 120)
private_college <- rnorm(100, mean = 1450, sd = 100)
community_college <- rnorm(100, mean = 1050, sd = 150)

# Combine into a single dataframe
admissions_data <- data.frame(
  college_type = factor(rep(c("Public", "Private", "Community"), each = 100)),
  sat_score = c(public_college, private_college, community_college)
)

# Ensure SAT scores are within realistic range (400-1600)
admissions_data$sat_score <- pmin(pmax(admissions_data$sat_score, 400), 1600)

# Quick look at the data
head(admissions_data)
```

## Task 2: Calculate Basic Statistics

```{r summary-stats}
# Summary statistics by college type
admission_summary <- admissions_data |>
  group_by(college_type) |>
  summarize(
    mean_score = mean(sat_score),
    median_score = median(sat_score),
    sd_score = sd(sat_score),
    min_score = min(sat_score),
    max_score = max(sat_score)
  )

print(admission_summary)
```

**Task 3** Describe the data

Write a couple of sentences describing the differences in SAT scores between types of colleges, using the descriptive statistics we just generated.

It seems that public and private schools admit people with significantly higher SAT scores than community colleges. The private schools get the most high scoring applicants out of the two. I'm not sure what the sd score signifies? I'm kind of hung up on how the standard deviation isn't a percent I suppose?

## Task 4: Visualize the Data

Let's create visualizations to see the patterns:

```{r visualize-distributions}
# Create density plot showing overlapping distributions
ggplot(admissions_data, aes(x = sat_score, fill = college_type)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "SAT Score Distributions by College Type",
    subtitle = "Are these differences statistically significant?",
    x = "SAT Score (400-1600 scale)",
    y = "Density",
    fill = "College Type"
  ) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1")
```

```{r boxplot}
# Create boxplot for another view
ggplot(admissions_data, aes(x = college_type, y = sat_score, fill = college_type)) +
  geom_boxplot() +
  labs(
    title = "SAT Scores by College Type",
    x = "College Type",
    y = "SAT Score"
  ) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1")
```

## Task 5: Perform ANOVA Test

```{r anova-test}
# Run one-way ANOVA
anova_result <- aov(sat_score ~ college_type, data = admissions_data)
summary(anova_result)
```

The F-value is huge (284.3). The F-value measures the ratio of between-group variance to within-group variance, and this high value suggests the differences between college types far outweigh the variation within each type. The p-value is very low, indicating the differences between college types are statistically significant at any reasonable significance level. The three asterisks (***) highlight that this is significant at the 0.001 level. So we can say that with confidence that there are real differences in SAT scores between the three types of colleges.

## Task 5: Compare the differences between specific pairs

```{r post-hoc}
# Tukey's HSD test to see which specific pairs of college types differ
tukey_result <- TukeyHSD(anova_result)
print(tukey_result)
```

The Tukey multiple comparisons test shows highly significant differences in SAT scores between all college types:

* Private vs. Community: Private colleges have SAT scores that are 366.96 points higher than community colleges on average (p < 0.0001)
* Public vs. Community: Public colleges have SAT scores that are 292.55 points higher than community colleges on average (p < 0.0001)
* Public vs. Private: Public colleges have SAT scores that are 74.40 points lower than private colleges on average (p = 0.000021)

All comparisons show extremely small p-values, indicating these differences are statistically significant at any conventional significance level. The 95% confidence intervals (shown in the "lwr" and "upr" columns) do not include zero for any comparison, further confirming these differences are statistically significant.

## Task 6: Report the Findings

Based on our analysis, we can now answer:

1. Is there a significant difference in SAT scores across college types?
2. Which specific college types differ from each other?
3. How large are these differences?

Answer those questions, providing evidence for your answers, and then write a headline and a lead paragraph reporting these findings using active voice and focusing on the differences across types.

1. There is a very significant difference in SAT scores across college types. We know this from a statistical point of view because the F value is very high which means that there's much variance between the groups from my understanding. The P-value is also very low because it was determined through the code that there was little coincidence involved essentially.

2.The two types that differ from eachother the most are the community colleges and the private/public colleges. There is obviusly a disperity between the two statistically because the private and public colleges admit the students with the generally higher SAT scores while the community colleges admit the lower scoring students. 

3. The differences are high. There is a near 300 score difference between public collges and community colleges.

Are community colleges becoming more mainstream?

Its getting harder and harder to get into good colleges in America as admission rates are heading down. In a recent study. Conducted the average score of a community college student on the SAT was a 1060, while that of public and private college students were 1435 and 1360, which is significantly higher. Does this mean that as time goes on, community colleges will become easier, and possibly more efficient choice when it comes to education?
