Final Project Code Tests



```{r}
library(tidyverse)
```

```{r}
Texas-Govoners <- read_csv(Texas_Govoner_Elections_Sheet1)
```




```{r}
i_and_e <- read_csv("Immigration_and Emmigration_Information_Sheet_1_1.csv")
```


```{r}
library(tidyverse)

migration_data <- read_csv("imeminfo.csv")

migration_long <- migration_data |>
  pivot_longer(
    cols = -Year,
    names_to = "route",
    values_to = "people"
  ) |>
  separate(route, into = c("origin", "destination"), sep = "_to_") |>
  mutate(people = as.numeric(gsub(",", "", people)))
```

```{r}
outflow_data<-read.csv(Immigration_and_Emmigration_Information_Sheet1_2)

```
```{r}


tx_pres|>
  
ggplot(aes(x = year))+
  geom_histogram(binwidth = 0.02, fill= "blue", color= "orange") + 
  labs(title= "Democrat Vote Count in Presidential Elections in TX", x= "Year", y= "vote count")
  
fl_pres|>
  
ggplot(aes(x = year))+
  geom_histogram(binwidth = 0.02, fill= "blue", color= "orange") + 
  labs(title= "Democrat Vote Count in Presidential Elections in FL", x= "Year", y= "vote count")
  
  
tx_gov|>
  
ggplot(aes(x = year))+
  geom_histogram(binwidth = 0.02, fill= "blue", color= "orange") + 
  labs(title= "Democrat Vote Count in Gubernatorial Elections in TX", x= "Year", y= "vote count")
  
fl_gov|>
  
ggplot(aes(x = year))+
  geom_histogram(binwidth = 0.02, fill= "blue", color= "orange") + 
  labs(title= "Democrat Vote Count in Gubernatorial Elections in FL", x= "Year", y= "vote count")
```  
  
  
immigration_data|>

what I'm trying to figrue out here is how I can make 4 lines on a line graph to show all four paths being taken. SO one for each states migration to the other state. THe x would be the the year and the y would be the number of people leaving.

belopw is what AI spit out when i put it in.



library(ggplot2)

# Example data frame
df <- data.frame(
  year = rep(2011:2022, each = 4),
  migration_path = rep(c("CA_to_TX", "CA_to_FL", "NY_to_TX", "NY_to_FL"), times = 12),
  people_leaving = sample(8000:25000, 48, replace = TRUE) # Replace with your actual data
)

# Plot
ggplot(df, aes(x = year, y = people_leaving, color = migration_path)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  labs(title = "Migration Flows Between States",
       x = "Year",
       y = "Number of People Leaving",
       color = "Migration Path") +
  theme_minimal()
  
  
  ///
model <- lm(`` ~ ``, data = immigrationxvotingrecords)

summary(model)
  ///
  
  
  
  
  
  
  
  
  
  
  
  
