---
title: "Identifying Schools with Biggest Differences in Math Test Scores, Part 2"
author: "Max Cassett"
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
options(scipen = 999)
```

### Introduction

We'll do what we did in class, but with a twist: instead of Montgomery County, Pa., you'll be working with data from Montgomery County, Md. Same basic idea, though: linear regression looking at free and reduced-price meals and math test scores.

Just as in the exercise, you will create a scatterplot to visualize the relationship between the FARMS (Free and Reduced-price Meal System) percentage and the percentage of students who met or exceeded math expectations. You'll then use linear regression to identify the schools that show the biggest differences between their actual and expected math test scores and evaluate the results of your model.

### Loading the Data

First, let's load the data into a data frame using the read_csv() function from the tidyverse library.

```{r}
schools <- read_csv("https://raw.githubusercontent.com/dwillis/jour405_files/refs/heads/main/montgomery_md.csv")
```
### Task: Create a Scatterplot and Describe the Results (5 points)

First, let's create a scatterplot to visualize the relationship between the FARMS percentage and the percentage of students who met or exceeded math expectations. We'll use ggplot2 to create the plot and add a linear regression line.

```{r}
schools |>
  ggplot(aes(x = `FARMS Pct`, y = `% Met or Exceeded Math Expectations`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(x = "FARMS Percentage", y = "% Met or Exceeded Math Expectations", title = "Scatterplot of Math Expectations vs FARMS Percentage")
```
Based on what you see here, what's the relationship between FARMS percentage and math test scores? Compare this scatterplot to the one you made in the class activity.

The relationship seems to be that the higher the FARMS percentage is, the lower the percentage of students met or exceeded math expectations. This scatterplot is a lot more spread out, rather than a lot of schools congregating around the 25% mark. The scatterplot in the activity indicated that there was an upward trend or at least thatb a student was more likely to meet expectations or exceed them in math if their school was in the FARMS program. This scatter plot indicates the opposite. 

### Task: Linear Regression Model (7 points)

Next, you'll perform a linear regression analysis using the lm() function. We'll set the percentage of students who met or exceeded math expectations as the dependent variable, and the FARMS percentage as the independent variable. In other words, if we know the FARMS percentage for a school, how reliably can we predict its math score?

```{r}
model <- lm(`% Met or Exceeded Math Expectations` ~ `FARMS Pct`, data = schools)

summary(model)
```

Provide the model's output for the three statistics mentioned in the class activity.

The P-value is low ( 0.00000000000000022) which indicates that the findings are not statisitcally signigicant which is odd but I'll roll with it. The adjusted R squared is 0.6734 which is kind of normal and high at the same time. The R squared adjusted is the measure of how much of the difference between schools' math scores can be explained by their FARMS percentage, in this case, it can be said there is a correlation. Lastly the coefficients are that thing  never wanted to deal with again, good ole algebra. Granted, cant say what I was expecting, this is a statistics class. The intercept is 68.04878 and the slope (m) is -0.59511, We can use that to figure out where a school would fall on on the plot using one variable or the other.

Based on what you've learned so far, what does it tell us about the nature of the relationship between FARMS percentage and math scores? Is this model worth exploring further? Why or why not?

It tell us that there really isnt anything here worth looking at and there is no statistically significant relationship.

### Task: Identifying Schools with Biggest Residuals (3 points)

Let's calculate and examine the residuals next. 

```{r}
schools$predicted <- predict(model)

# Calculate the residuals
schools <- schools |> 
  mutate(residual = `% Met or Exceeded Math Expectations` - predicted)

# Sort the data frame by the absolute value of the residuals in descending order
schools <- schools |>
  mutate(abs_residual = abs(residual)) |>
  arrange(desc(abs_residual))

```

Look at the absolute values of the residuals; what are the largest differences between the predicted result and the actual math scores? Pick two schools with the largest residuals. Find the zip codes where they are located and using the CensusReporter.org website, describe the most interesting demographic data from each zip code.


The largest the absolute residuals go is 29.99392462 points and the smallest is near zero. So in some cases the predictions were totally off and in other they were spot on.

SO i went a little further and looked into the zip codes of the elementary schools on both sides of the spectrum, and I found some interesting results. First off there is a huge difference in the average income of those who are predicted to do worse in math. for instance in the first three elementary school the average income didnt break 55 thousand. While those elementary schools that were predicted to do well in math and the predictions panned out, were in 100k average income areas. So from my understanding, people are betting against those kids whose families are lower income. Not only that but there is a significant black of population in the first two schools that had large residuals. Although in the case of the third school with the third highest residuals, they were predicted to do better than they actiually did

## Task: Evaluate the model (5 points)

Describe what conclusions you'd draw from this model's results and how it would influence a story pitch about the school board member and math scores.

I think there is a story here about betting against impoverished people and how that might be reflected in funding. If people are predicting that kids from poor families arent going to do well in school then that might influence whether a city council wants to deviate funds to a better more successful school.Why is there such a low opinion of these students? Could it be racially motivated. there is a lot of good data to use here.