---
title: "HW11: Education Level and Survey Weighting"
author: "Your Name Here"
---

```{r setup, include=FALSE}
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
```

# Introduction

In this assignment, you'll examine how survey weighting affects our understanding of voting intentions across different education levels, using the data we examined in class. You'll calculate both weighted and unweighted statistics, create visualizations, and reflect on the implications for reporting.

## The Data

The nonvoters dataset contains survey responses about voting intentions and behaviors from a national survey. The survey was conducted prior to an election and includes demographic information like education level.

```{r}
# Load the dataset
nonvoters_data <- read_csv("https://raw.githubusercontent.com/dwillis/jour405_files/refs/heads/main/nonvoters_data.csv")

# Take a quick look at the data structure
glimpse(nonvoters_data)
```

### Key Variables

- `weight`: Survey weight assigned to each respondent
- `Q21`: Voting intention (1 = Yes, 2 = No, 3 = Unsure/Undecided)
- `educ`: Education level (College, Some college, High school or less)

## Task 1: Education Distribution

First, let's examine the distribution of education levels in our sample. Replace "REPLACE_ME" with the correct variable for education level.

```{r}

education_distribution <- nonvoters_data |>
  count(educ) |>
  mutate(percentage = n / sum(n) * 100) |>
  kable(digits = 1, col.names = c("Education Level", "Count", "Percentage (%)"))

education_distribution
```

## Task 2: Reflection Question

Why might education levels in survey samples often differ from the general population? What factors might cause certain education groups to be over or underrepresented?

They might differ because of where the survey took place. Like for instance there would be more college graduates voting that live in Bel Air in LA than San Marcos Texas. The demographics are important to take into account when surveying.


## Task 3: Unweighted Analysis by Education

Now, let's calculate unweighted voting intentions by education level. This is what we would report if we didn't apply any weighting to our sample.

```{r}
# Calculate unweighted voting intentions by education
unweighted_by_education <- nonvoters_data |>
  # Filter out missing values
  filter(!is.na(Q21), Q21 > 0, !is.na(educ)) |>
  # Group by education and response
  group_by(educ, Q21) |>
  # Count responses
  summarize(count = n(), .groups = "drop_last") |>
  # Calculate percentages
  mutate(total = sum(count),
         percentage = count / total * 100) |>
  ungroup()

# Create a more readable format with voting intentions as columns
unweighted_educ_summary <- unweighted_by_education |>
  pivot_wider(
    id_cols = educ,
    names_from = Q21,
    values_from = percentage,
    names_prefix = "pct_"
  ) |>
  rename(
    "Yes (%)" = pct_1,
    "No (%)" = pct_2,
    "Unsure (%)" = pct_3
  )

kable(unweighted_educ_summary, digits = 1, caption = "Unweighted Voting Intentions by Education Level")
```

## Task 4: Reflection Question

Based just on this unweighted analysis, what headline might you write for a news story about education and voting intentions?

I'd probably beg the question of why people with less education vote. Do people who have less of an education distrust the worth of their vote more?

## Task 5: Weighted Analysis by Education

Next, let's apply survey weights to see how this changes our results. Instead of just counting responses, we'll sum the weights for each group. Replace "REPLACE_ME" with the appropriate weight variable

```{r weighted-by-education}

weighted_by_education <- nonvoters_data |>
  # Filter out missing values
  filter(!is.na(Q21), Q21 > 0, !is.na(educ)) |>
  # Group by education and response
  group_by(educ, Q21) |>
  # Sum the weights instead of counting
  summarize(weighted_count = sum(weight), .groups = "drop_last") |>
  # Calculate weighted percentages
  mutate(weighted_total = sum(weighted_count),
         weighted_percentage = weighted_count / weighted_total * 100) |>
  ungroup()

# Create a more readable format
weighted_educ_summary <- weighted_by_education |>
  pivot_wider(
    id_cols = educ,
    names_from = Q21,
    values_from = weighted_percentage,
    names_prefix = "pct_"
  ) |>
  rename(
    "Yes (%)" = pct_1,
    "No (%)" = pct_2,
    "Unsure (%)" = pct_3
  )

kable(weighted_educ_summary, digits = 1, caption = "Weighted Voting Intentions by Education Level")
```

## Task 6: Reflection Questions

1. How did the percentages change after applying weights? Which education group showed the biggest changes?

There was a loss of percent of those who only had a high school education or less voting and naturally a higher percentage of those who said No in regards to voting with a high school education or less because there was a loss.No major changes otherwise.

2. Why might the weighted results be considered more accurate than the unweighted results?

I suppose they are more accurate because the weight negates the screw that the data takes when one group is over represented.

## Task 7: Comparison of Weighted vs. Unweighted Results

Let's create a direct comparison table to see the differences more clearly.

```{r}

comparison <- unweighted_educ_summary |>
  inner_join(weighted_educ_summary, by = "educ", suffix = c("_unweighted", "_weighted")) |>
  mutate(
    # Calculate the differences between weighted and unweighted percentages
    yes_diff = `Yes (%)_weighted` - `Yes (%)_unweighted`,
    no_diff = `No (%)_weighted` - `No (%)_unweighted`,
    unsure_diff = `Unsure (%)_weighted` - `Unsure (%)_unweighted`
  ) |>
  # Select just the columns we want to display
  select(educ, yes_diff, no_diff, unsure_diff) |>
  rename(
    "Education Level" = educ,
    "Yes (% point diff)" = yes_diff,
    "No (% point diff)" = no_diff,
    "Unsure (% point diff)" = unsure_diff
  )

kable(comparison, digits = 1, caption = "Difference Between Weighted and Unweighted Results (percentage points)")
```

## Task 8: Reflection Question

Which education group shows the largest differences between weighted and unweighted results?

The high school education group 

## Task 9: Visualization

Visualizations can help us see the differences more clearly. Let's create a bar chart comparing weighted and unweighted "Yes" responses by education level. Replace "REPLACE_ME" with the correct variable name

```{r}
educ_viz_data <- bind_rows(
  # Unweighted data
  unweighted_by_education |> 
    filter(Q21 == 1) |>  # Only "Yes" responses (Q21=1)
    mutate(Type = "Unweighted") |>
    select(Type, educ, percentage),
  
  # Weighted data - 
  weighted_by_education |> 
    filter(Q21 == 1) |>  # Only "Yes" responses
    mutate(
      Type = "Weighted",
      percentage = weighted_percentage
    ) |>
    select(Type, educ, percentage)
)

# Create a grouped bar chart
ggplot(educ_viz_data, 
       aes(x = educ, y = percentage, fill = Type)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = sprintf("%.1f%%", percentage)), 
            position = position_dodge(width = 0.9), vjust = -0.5) +
  labs(
    title = "Weighted vs. Unweighted 'Yes' Responses by Education",
    subtitle = "Q21: Do you plan to vote in the November election?",
    y = "Percentage (%)",
    x = "Education Level"
  ) +
  scale_fill_manual(values = c("Unweighted" = "#619CFF", "Weighted" = "#F8766D")) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold"),
    legend.position = "bottom"
  )
```

## Task 10: Reflection Questions

Does the visualization make it easier to see the differences between weighted and unweighted results? Why or why not?

Yeah I mean it quite literally shows the differences visually so yes in a way. But it lookes like it shows that there really isnt much of a difference which means that the survey is pretty good at avoiding any survey bias.

## Task 11: Summary

Based on your analysis of weighted and unweighted results by education level, write a brief (2-3 paragraph) journalistic summary of what you found. You should try to address:

1. How education level relates to voting intentions
2. How weighting affected your understanding of this relationship
3. What this means for interpreting polling results in news reporting


  The surveys show that people with a lower education level intend to vote less than those with a higher level of education. This can have its own consequences in the world such as thos of lower education having less of a say in who runs our nation, which may or may not be a good thing. Probably is a good thing that the people who are most educated and informed are making those decisions but thats besides the point.
  Weighting is simply a tool used to beat the skew that data takes when a certain people group is over represented in the survey. If i'm being frank I'm not sure i complelty understand the significance of a few percentile points but it does help give a more accurate representaiton from the data.



## Task 12: Final Reflection Questions

1. Why is it important for journalists to understand survey weighting when reporting on polls?

Because it effects the data in away that is more nuanced then statistics. It gives a more accurate depitication of the data by negating it in a way.

2. How might the differences between weighted and unweighted results affect how you would report on this data?

I am confused by how much weight is being pu on weighting. It seems to only be small cahnges in the weight, like by a few percent. I suppose that if there a greater number of a certain populace surveyed that was unequal to other groups then it would be important to weight the results so that you had a more accurate report.

3. What additional information would you want to know about how the weights were calculated before using this data in a news story?

Where this survey took place, family income, race, gender, a slew of other things about the individuals so the data could be weighted further to create an even more accurate picture.
