---
title: "HW2 Prince George's County Crime"
---

In this assignment, we will be working with a dataset reporting the number of crimes by month in Prince George's County, Maryland, for every month in 2024.

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: <https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv>. You will create a dataframe called `pg_crime` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
pg_crime<-read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv")
```

### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, are any of the categories more important than others? Why or why not? Do any of them make the job of summarizing the data more difficult? Why or why not?

It looks like there is a ton of automotive crime in PG county. The amount of cars being broken into, stolen, carjacked, etc, its ridiculously high. Regular theft doesnt even break the top four at any point during the year, not to say that theft isnt high though. As a journlaist reading this I would ask a few questions; whats being done to prevent these crimes, and what empowers these criminals to be so bold to steal cars? I'd ask how many of these auto thefts end with the car being returned and what part do traffic cameras play on that? Where are the majority of these accidents and autothefts happening? Is there a tend there. I think that the accident numbers are important to. Where are these accidneets happening the most and such are the questions I would tack on to this data. I think that the dangerous crimes are definelty the more important to get answers on in this data set because getting answers and holing people accountable to get solutions would help save lives and livlihoods. I think that this data set is a little unclear with the collumn to the very left and the numbers. 

### Task 4: Calculating Averages

Calculate the total number of incidents per month for each category using `mutate()`, calling the new column `total_incidents`. Then calculate the average number of incidents per month for each category of crime, calling the new column `average_incidents`. Assign the result to a new dataframe called `pg_crime_averages`.

Ok so I had to get some help for this one but I'm going to write out my explanation so I know I understand whats happening here. Pg_crime is getting a new column (total incidents) with the |>, and the column is mutating the information by taking the total of all the data for a certain crime and adding that up. in the second line the total number of incidents is taken and aveeraged over four and the |> creates that column. In the last line of code the name of the chart is given.

```{r}
total_incidents<- pg_crime|> mutate(total = Jan+Feb+Mar+Apr+May+Jun+Jul+Aug+Sep+Oct+Nov+Dec) 
average_incidents <- total_incidents |> mutate(average = total/12)
pg_crime_average_crime_incidents <- average_incidents 
```

### Task 5: Describe the Averages

Examine the new dataframe, `pg_crime_averages`, by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the average number of incidents per month for each category. Which categories have the highest average number of incidents per month? Which have the lowest? What does this tell you about the pattern of crime in Prince George's County? Are there categories that you have questions about? Why or why not?

Like I mentioned before the auot related crimes are the highest reported ones so it makes sense that they have the higer averages. Theft doesnt break the top four again. Something that jumped out at me was the lack of violent crimes in Pg county as a whole. The jump from assault to theft is hefty. A question I have is if better public transportaiotn would help lower the auto crim rate because people would eliminate the need for a second or third car because of their use of a bus or tram on a regular basis. 


### Task 6: Testing a Theory

One theory about crime is that it is more likely to occur in the summer months. Test this theory by creating three new columns using `mutate()` that calculate the percentage change compared to the average for each category of crime for the months of June, July, and August. Call the new columns `june_change`, `july_change`, and `august_change`. Assign the result to a new dataframe called `summer_crime_averages`.


I hate this, this is nuts, the fact that people understand this blows my mind. I want to understand so bad. What meeting did I miss where this became common knowledge. I'm going to wake steve jobs from the dead and get him to help me out so help me lord.

```{r}
june_change <- pg_crime|> mutate(june_change = Jun-May/May*100) 
july_change <- pg_crime|> mutate(july_change = Jul-Jun/Jun*100) 
august_change <- pg_crime|> mutate(august_change = Aug-Jul/Jul*100) 
summer_crime_averages <- pg_crime |> mutate((august_change+ july_change+june_change)/3)

```


### Task 7: Describe the Summer Crime Averages

Look at the `summer_crime_averages` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change in the average number of incidents per month for each category of crime for the months of June, July, and August. Do you see any patterns? What does this tell you about the theory that crime is more likely to occur in the summer months?

unfortunatley I couldnt figure this one out. I tried to get it all in the same chart but even when i looked it up I coudnt understand. From looking at the chart I think that there is more crime towards the end of the year than the start and the summer doesnt have a correlation with higher crime.

### Task 8: What's the Story?

Write up a lede and nutgraf of a story focused on the crime data in Prince George's County. It could be from any of the dataframes you've created in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

I'm going to say that the autotheft story is the one i would go with. my lede would be

PG County's auto crimes out do the number of robberies, assaults and sex crimes, but what makes lifting a car so appealing?

Residents in PG county might be teased on the dangerous reputation of their community but the truth might suprise you. The amount of crimes involving motor vehicles far outweigh those that do not. Crimes that pertain to accidents on the road way take fourth and first place, but between the two with similar numbers are auto theft and theft from a vehicle. What are stragedies that criminals use to rob you of your vehicle and from theirs? What emboldens them to do so and why are they so successful? more on that below.


### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
