---
title: "HW3 WH Salaries"
---

In this assignment, we will be working with a dataset containing information about the salaries of White House staffers as of July 2024.

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/biden_wh_salaries_2024.csv. You will create a dataframe called `wh_salaries` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
wh_salaries <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/biden_wh_salaries_2024.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, what stands out to you about this data? Are there any salary values that will cause problems for analysis?

I noticed that out of all of his advisors on policy one of his immigration advisor is paid the most which means he Fonald Trump values their experteace as important above all other advisors, which tell us about what his most pressing plans are for the country. The second highest paid was a native affairs policy advisor followed by workforce policy and criminal justice. I assume that with the current climate around tarrifs trump would want to focus pretty heavily on workforce policy, so it serves to know who this guy is, what his plan is, and why is he paying him 180,000 dollars

### Task 4: Calculate the Mean and Median of the Salaries

Write code to calculate the mean and median of the salaries in the dataset using the `summarize()` function. You don't need to make a new dataframe, just display the results.


I had to use perplexity AI for this opne. I essentually asked it the question and told it the data set and that the code was in R. Is there a place that we're shown this code that I'm not aware of or that I missed because I'm feeling really lost and I find myself guessing and in the dark when it comes to writing this code. I'm not sure that I understand what this code is doing either. 

```{r}
wh_salaries |>
  summarize(
    mean_salary = mean(salary, na.rm = TRUE),
    median_salary = median(salary, na.rm = TRUE)
  )
```

### Task 5: Add in the minimum and maximum salaries

Take the code you just wrote and add in the minimum and maximum salaries (the functions for that are `min()` and `max()`). Your results should show four columns: mean_salary, median_salary, min_salary, max_salary.

for this one my first try I essentially copied and pasted my previuos answer like so;

wh_salaries |>
  summarize(
    mean_salary = mean(salary, na.rm = TRUE),
    median_salary = median(salary, na.rm = TRUE)
    min_salary = min(salary, na.rm = TRUE)
    max_salary= max(salary, na.rm = TRUE)
    )
    
    Unfortunaly that didnt work and I wish it didnt because that would mean that I understood something. i assume I'm on the write track as far as columns go. all i had to do was add commas, thats insane that a comma took away 15 minutes of my life. 

```{r}
wh_salaries |>
  summarize(
    mean_salary = mean(salary, na.rm = TRUE),
    median_salary = median(salary, na.rm = TRUE),
    min_salary = min(salary, na.rm = TRUE),
    max_salary= max(salary, na.rm = TRUE)
    
  )
```

```

### Task 6: Describe the Results

Write a couple of sentences describing the results of your analysis. What do the mean and median salaries tell you about the data? What do the minimum and maximum salaries tell you? Are there any outliers in the data?

Knowing the mean and the median are important because it shows the average wage thats being payed in the white house as well as where the amount skews as far as lower or higher than the average. there are repeating zeros and such but I wouldnt consider those outliers in the data.  


### Task 7: Story Idea

Based on the data, what is your best story idea? What is the next step you would take to develop that story idea?


My best story idea is a basic story on whos getting paid in the top bracket at the white house and what their jobs entail and how their jobs, epxertise and history give us a clue into trumps future endeavors.

