---
title: "HW6_Spotify_Songs"
---

In this assignment, you will be working with a dataset of songs played on Spotify in 2020, and you'll be writing code to explore the range of danceability scores in the data, using standard deviation and histograms. You'll also write up your findings.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)

```

### Task 2: Load the Spotify Songs dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/spotify_2020.csv

```{r}
spotify_songs<- read.csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/spotify_2020.csv")

```

### Task 3: Calculate the mean and standard deviation of the danceability scores

Write code to calculate the mean standard deviation of the danceability scores in the Spotify dataset. Display those.

```{r}
spotify_songs <- spotify_songs %>% 
  mutate(danceability_mean = danceability_mean, danceability_sd = danceability_sd)


danceability_mean <- mean(spotify_songs$danceability, na.rm = TRUE) 
danceability_sd <- sd(spotify_songs$danceability, na.rm = TRUE) 


spotify_songs |>
  summarize(danceability_sd = sd(danceability), danceability_mean=mean(danceability))
"this was the code form the review"
```

Used chat gpt but I think that this was actually a great decision because I feel like I'm finally understanding the langauge better. Mutating spotify songs and seeing that happen was pretty eyeopening ngl. As far as the code goes from the mean and standard deviation its pretty basic.

### Task 4: Create a histogram of the danceability scores

Write code to create a histogram of the danceability scores in the Spotify dataset, based on the example we did in class.

```{r}
spotify_songs|>
  
ggplot(aes(x = danceability))+
  geom_histogram(binwidth = 0.02, fill= "blue", color= "orange") + 
  labs(title= "distribution of Dancability Scores in Spotify Songs 2020", x= "danceability score", y= "number of songs")


```

### Task 5: Add a line to the histogram for the mean danceability score

Write code to add a vertical line to the histogram for the mean danceability score, based on the example in class.

```{r}
spotify_songs|>
ggplot(aes(x = danceability))+
  geom_histogram(binwidth = 0.02, fill= "blue", color= "orange") + 
  geom_vline(aes(xintercept = danceability_mean), color = "red", linetype = "dashed", size = 1) +
  labs(title= "distribution of Dancability Scores in Spotify Songs 2020", x= "danceability score", y= "number of songs")

```

### Task 6: Describe our results

Write a few sentences about the distribution of danceability scores in the Spotify dataset. Is it a normal distribution, or is it skewed in one direction? What does the mean and distribution tell us about the songs in the dataset?

It looks to me that the danceablilty of the selected songs in the data set is fair highly, being just below 70% dance able. It doesnt look like there is a certain skew other than obviuosly  the majority of songs were dancable. 

### Task 7: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.

